Assessing \Htwo{} and \COtwo{} Network Strategies in Europe

### 2023-12-13 report/submissions/journal/paper/paper.tex
workflow: write out plotted data to csv files
update numbers to new results adjust legend in capture line plot

Total annual system costs in the \modBase{} are highest, totalling \label{}803 billion euros. The models with additional transport options for carbon and hydrogen achieve a more efficient system allocation and lower total system costs, with around \label{}2\% cost reduction (\label{}781 billion euros) in the \modCO{}, and around \label{}4\% cost reduction in the \modH{} and the \modHybrid{} (\label{}765 and \label{}764 billion euros, respectively).

% TODO: avoid to long paranthesis
In all models, almost half of the system costs are spent on primary electricity production from wind and solar power (\label{}45\% or \label{}360 of \label{}799 billion euros total costs in the \modBase{}), with another \label{}12\% (\label{}100 billion euros) spent on hydro, nuclear and biomass. Costs for energy transport networks for electricity, gas, hydrogen and carbon are below \label{}12\% (\label{}100 billion euros), with the electricity grid making up more of the system costs than the other three networks combined. This is the case for all four models. Capital expenditures on carbonaceous fuel production are around \label{}3\% (\label{}25 billion euros) in each model, while electrolysis makes up around \label{}5\% (\label{}40 billion euros). Heat pump installations account for around \label{}8\% (60 billion euros) of total system costs.

@@

\subsection*{Transport systems unlock point\=/source carbon capture}

In comparison to the other three models, the \modBase{} reveals significant inefficiencies: On the one hand, regions with no access to sequestration sites (i.e., ``landlocked'' regions) underutilize CC potentials from process emissions, fossil gas and biomass. On the other hand, sequestration regions rely on costly DAC on-site. Furthermore, hydrogen and power prices show high regional differences, indicating system inefficiencies due to transport constraints caused by non\=/existent hydrogen or insufficient power transmission. Prices are significantly higher in Central Europe, a region with high population density and high concentration of energy-intensive industries.

By enabling the efficient transportation of carbon, hydrogen, or both, these models reduce the reliance on high-cost DAC, and source more carbon from low-cost point-source CC on biomass and gas in industrial applications (see Fig.~\ref{fig:balance_captured_carbon}). The \modH{} only captures \label{}200~Mt of \COtwo{} from DAC compared to \label{}400~Mt in the \modBase{}. Instead, the \modH{} captures more carbon from process emissions as well as gas and biomass for industry. The \modCO{} uses only \label{}150~Mt from DAC, and instead also captures more carbon than the \modH{} from gas for industry and biomass CHPs. And finally, the \modHybrid{} captures only \label{}100~Mt from DAC, but slightly more carbon from all other CC technologies.
Also, the networks reduce the reliance on methanation in the model, reducing the overall demand for CC volumes. Another knock-on effect of the reduced reliance on DAC is reduced endogenous demand for renewables to power DAC facilities.
% TODO: make sure FT and methanolisation is properly explained in methods section as almost exogenously fixed products

@@

In the \modBase{}, the capture share is highest for biogas-to-gas facilities, followed by process emissions, biomass for industry, biomass CHPs, gas for industry, SMR and gas CHPs. For the network models, we can observe a correlation between CC share and capacity factors of the underlying technologies: Process emissions as well as gas and biomass for industry have high capacity factors (above 80\%), and CC shares close to 100\%, the only exception being gas for industry in the \modH{}. Biomass CHPs have a capacity factor of around 40\%, and CC shares of 50\% in the \modH{} and 95\% in the \modCO{}. SMR and gas CHPs have capacity factors below 20\%, and CC shares around 50\% and 0\%, respectively, across all network models. Gas CHPs serve as peak load electricity production technology, and with very few operating hours, thus the high investment costs into CC applications are not efficient. Similarly, SMR serves as ``peak-load'' hydrogen production technology in regions with poor renewable resources for electrolysis.

@@

In transitioning from a net zero to a net emissions reduction target, a continuation of investment trends are perceived. Fig.~\ref{fig:net-negative_cost_bar} shows the net investment changes per technology group for the \Hgrid{} and \modHybrid{}s. The scenario imposes a net carbon removal of 460 Mt/a, with both models incurring a similar net cost increase - 95 billion €/a for the \modH{} and 96 billion €/a for the \modHybrid{}, balancing the earlier cost advantage of the \modHybrid{}. Both models invest relatively evenly in additional DAC, wind and heat pump systems, with DAC accounting for more than a third of the additional system costs. These remove carbon directly at sequestration sites at the coast, in particular in Portugal and UK, together with heat and power from new heat pumps at site and new solar and wind installation in the vicinity.
More heat pumps, solar and wind installations across all Central and Eastern European countries as well as biogas-to-gas installations compensate for the reduction of fossil gas based heating and powering. Reduced usage of gas in general lower the investments in gas infrastructure in both models.
Fundamental difference in investment strategies occur in the carbon capture sector: In contrast to the \modH{}, the \modHybrid{} continues the strategy of sequestering carbon captured and transported from bio-energetic inputs, expanding the carbon transport routes from Central and Eastern Europe to the near shores. At the same time, carbonaceous fuel production in Central Europe partially moves to Spain and UK, where it uses carbon from new DAC facilities without the installation of a \COtwo{} network. At the same time, solar based electrolysis in France, Nothern Italy and Switzerland expands requiring more solar capacities but less hydrogen imports from the Iberian Peninsula.
The \modH{} on the other hand, expands decentralized CU systems from the Net-Zero scenario, which are now increasingly supplied with carbon from biogas-to-gas facilities, showing up in both ``Bioenergy'' and ``Carbon Capture at Point Sources'' cost-contributions. The centralized DAC facilities at sequestration sites are more expanded than in the \modHybrid{} and are requiring additional heat from biomass CHPs.

>>>>

Total annual system costs in the \modBase{} are highest, totalling \label{}779 billion euros. The models with additional transport options for carbon and hydrogen achieve a more efficient system allocation and lower total system costs, with \label{}3.1\% cost reduction (\label{}754~bn€) in the \modCO{}, and \label{}4.8\% (\label{}741~bn€) in the \modH{}, and \label{}5.2\% (\label{}738~bn€) in the the \modHybrid{}.

% TODO: avoid to long paranthesis
In all models, almost half of the system costs are spent on primary electricity production from wind and solar power (\label{}47\% or \label{}371~ billion euros in the \modBase{}), with another \label{}10\% (\label{}79 billion euros) spent on hydro, nuclear and biomass. Costs for energy transport networks for electricity, gas, hydrogen and carbon are below \label{}4\% (\label{}25 billion euros), with the electricity grid making up more of the system costs than the other three networks combined. This is not the case for the other models, where combined pipeline costs slightly surpass electricity transmission system costs (up to 20\% for the \hybridmodel). Capital expenditures on carbonaceous fuel production are around \label{}4\% (\label{}30 billion euros) in each model, while electrolysis makes up around \label{}4.7\% (\label{}37 billion euros). Heat pump installations account for around \label{}8\% (63 billion euros) of total system costs.

@@

\subsection*{Transport systems unlock point\=/source carbon capture}
i
In comparison to the other three models, the \modBase{} reveals significant inefficiencies: On the one hand, regions with no access to sequestration sites (i.e., ``landlocked'' regions) underutilize CC potentials from process emissions, fossil gas and biomass. On the other hand, sequestration regions rely on costly DAC on-site. Furthermore, hydrogen and power prices show high regional differences, indicating system inefficiencies due to transport constraints caused by non\=/existent hydrogen or insufficient power transmission. Prices are significantly higher in Central Europe, a region with high population density and high concentration of energy-intensive industries.

By enabling the efficient transportation of carbon, hydrogen, or both, these models reduce the reliance on high-cost DAC, and source more carbon from low-cost point-source CC on biomass and gas in industrial applications (see Fig.~\ref{fig:balance_captured_carbon}). The \modH{} only captures \label{}193~Mt of \COtwo{} from DAC compared to \label{}392~Mt in the \modBase{}. Instead, the \modH{} captures more carbon from process emissions as well as gas and biomass for industry. The \modCO{} uses only \label{}154~Mt from DAC, and instead also captures more carbon than the \modH{} from gas for industry and biomass CHPs. And finally, the \modHybrid{} captures only \label{}112~Mt from DAC, but slightly more carbon from all other CC technologies.
Also, the networks reduce the reliance on methanation in the model, reducing the overall demand for CC volumes. Another knock-on effect of the reduced reliance on DAC is reduced endogenous demand for renewables to power DAC facilities.
% TODO: make sure FT and methanolisation is properly explained in methods section as almost exogenously fixed products

@@

In the \modBase{}, the capture share is highest for biogas-to-gas facilities, followed by process emissions, biomass for industry, biomass CHPs, gas for industry, SMR and gas CHPs. For the network models, we can observe a correlation between CC share and capacity factors of the underlying technologies: Process emissions as well as gas and biomass for industry have high capacity factors (above 80\%), and CC shares close to 100\%, the only exception being gas for industry in the \modH{}. Biomass CHPs have a capacity factor of around 50\%, and CC shares of 50\% in the \modH{} and 97\% in the \modCO{}. SMR and gas CHPs have capacity factors around 15\%, and CC shares at nearly 0\% for the \baelinemodel. The capacity factor and CC share of SMR increases to around 20\% and 50-65\%, respectively, in the other models. Gas CHPs serve as peak load electricity production technology, and with very few operating hours, thus the high investment costs into CC applications are not efficient. Similarly, SMR serves as ``peak-load'' hydrogen production technology in regions with poor renewable resources for electrolysis.

@@

In transitioning from a net zero to a net emissions reduction target, a continuation of investment trends are perceived. Fig.~\ref{fig:net-negative_cost_bar} shows the net investment changes per technology group for the \carbonmodel, the \Hgrid{} and \modHybrid{}s. The scenario imposes a net carbon removal of 460 Mt/a, with all models incurring a roughly similar net cost increase - 77 billion €/a for the \modCO{}, 81 billion €/a for the \modCO{}, and 78 billion €/a for the \modHybrid{}, balancing the earlier cost advantage of the \modHybrid{}. Both models invest relatively evenly in additional DAC, wind and heat pump systems, with DAC accounting for more than a third of the additional system costs. These remove carbon directly at sequestration sites at the coast, in particular in Portugal and UK, together with heat and power from new heat pumps at site and new solar and wind installation in the vicinity.
More heat pumps, solar and wind installations across all Central and Eastern European countries as well as biogas-to-gas installations compensate for the reduction of fossil gas based heating and powering. Reduced usage of gas in general lower the investments in gas infrastructure in both models.
Fundamental difference in investment strategies occur in the carbon capture sector: In contrast to the \modH{}, the \modHybrid{} continues the strategy of sequestering carbon captured and transported from bio-energetic inputs, expanding the carbon transport routes from Central and Eastern Europe to the near shores. At the same time, carbonaceous fuel production in Central Europe partially moves to Spain and UK, where it uses carbon from new DAC facilities without the installation of a \COtwo{} network. At the same time, solar based electrolysis in France, Nothern Italy and Switzerland expands requiring more solar capacities but less hydrogen imports from the Iberian Peninsula.
The \modH{} on the other hand, expands decentralized CU systems from the Net-Zero scenario, which are now increasingly supplied with carbon from biogas-to-gas facilities, showing up in both ``Bioenergy'' and ``Carbon Capture at Point Sources'' cost-contributions. The centralized DAC facilities at sequestration sites are more expanded than in the \modHybrid{} and are requiring additional heat from biomass CHPs.
